Verifying Spatial Properties of Array Computations

### 2017-04-18 oopsla17/evaluation.tex
verification case studies

@@

\textbf{All pointed} \num{\overalljustPointed} of the stencils we
found involved only pointed regions.
% 39,681 of these were pointed in all dimensions.
Common examples of this were pointwise transformations on data (such
as scaling).

\textbf{Single-action} specifications comprise one
forward, backward, or centered region constant combined via \term{+}
or \term{*} with any number of pointed regions. We identified \num{\overallsingleAction} single-action
specifications, of which \num{\overallsingleActionIrr} were single-action with a
\texttt{nonpointed} modifier.

\textbf{Multi-action} specifications comprise at least two
 forward, backward, or centered regions, combined with
any number of pointed regions. We identified \num{\overallmultiAction} multi-action
specifications out of which \num{\overallmultiActionMulOnly} had regions combined only with
$\term{*}$ and \num{\overallmultiActionPlusAndMul} combined with a mix of $\term{*}$ and \term{+}.

@@

>>>>

\newcommand\regname[1]{\texttt{#1}}
\newcommand\pointed{\regname{pointed}}
\newcommand\forward{\regname{forward}}

@@

\textbf{All pointed} \num{\overalljustPointed} of the stencils we
found involved only \pointed\ regions.
% 39,681 of these were pointed in all dimensions.
Common examples of this were pointwise transformations on data (such
as scaling).

\textbf{Single-action} specifications comprise one
forward, backward, or centered region constant combined via \term{+}
or \term{*} with any number of \pointed\ regions. We identified \num{\overallsingleAction} single-action
specifications, of which \num{\overallsingleActionIrr} were single-action with a
\texttt{nonpointed} modifier.

\textbf{Multi-action} specifications comprise at least two
 forward, backward, or centered regions, combined with
any number of \pointed\ regions. We identified \num{\overallmultiAction} multi-action
specifications out of which \num{\overallmultiActionMulOnly} had regions combined only with
$\term{*}$ and \num{\overallmultiActionPlusAndMul} combined with a mix of $\term{*}$ and \term{+}.

@@

\subsection{Verification case studies}
\label{sec:case-studies}

\subsubsection{Catching bugs}

One problem we face while examining the millions of line of source
code is that we do not have the original programmer's intention at
hand. But sometimes we have the next best thing, the log entry in the
revision control system. Scanning through the UM we identified an
example of an off-by-1 error that could have been avoided through the
use of stencil specification. The buggy code was identified as
\pointed\ in all dimensions:

\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  !=stencil readOnce, (pointed(dim=1))*(pointed(dim=2))*(pointed(dim=3))
\end{minted}

But the programmer made it clear in the comments after the bug fix
that the intention had been \forward\ in the third dimension:

\begin{minted}[breakafter=+:,breakindent=-0.6em,breaksymbolsep=0.4em,linenos=false,xleftmargin=-0.5em]{fortran}
  !=stencil readOnce, (forward(depth=1, dim=3, nonpointed))*(pointed(dim=1))*(pointed(dim=2))
\end{minted}

\subsubsection{Aiding refactoring}

One of the initial goals of CamFort was to provide tools that enable
optimisation and refactoring without changing behaviour, and the
stencil specification feature was designed with that in mind. But
while we were perusing real-world source bases, we realised that there
were also many cases where CamFort could help with refactoring
precisely because they required changes in behaviour. For example, a
relatively common change observed in the logs of revision control for
various projects is the refactoring of array indices. Either
re-ordering, adding or deleting indices. Any of these, if not
perfectly propagated throughout all uses, could result in unexpected
outcomes. However, CamFort will pick up the difference during its
stencil specification checking. Ideally, an experienced programmer may
take advantage of the region variable feature of CamFort in order to
minimise the number of specifications that need changing and gain the
full time-saving advantage. But even if not, a quick search for the
specifications can have them updated quickly and then any code that is
not updated will trigger an error when checked.

% \subsubsection{Stress-testing CamFort}
